brados: declarative, programmable object storage

### 2016-05-25 paper.tex
some edits
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

\lstset{frame=tb,
  language=Ruby,
  basicstyle={\scriptsize\sffamily},
  breaklines=true,
  captionpos=b
}

%  aboveskip=3mm,
%  belowskip=3mm,
%  showstringspaces=false,
%  %columns=flexible,
%  basicstyle={\scriptsize\ttfamily},
%  numbers=none,
%  %numberstyle=\tiny\color{gray},
%  %keywordstyle=\color{blue},
%  %commentstyle=\color{dkgreen},
%  %stringstyle=\color{mauve},
%  %breaklines=true,
%  breakatwhitespace=true,
%  tabsize=3
%}

@@

\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
state do
  table :epoch, [:epoch]
  table :log, [:pos] => [:state, :data]

  interface input,  :op,   [:type, :pos, :epoch] => [:data]
  interface output, :ret,  [:type, :pos, :epoch] => [:retval]

@@

  # op's position was or was not found in the log
  scratch :found_op,    op.schema
  scratch :notfound_op, op.schema
end

bloom do
  # op's position found in log
  found_op <= (valid_op * log).lefts(pos => pos)
  notfound_op <= valid_op.notin(found_op)

  # demux on operation type
  write_op <= valid_op { |o| o if o.type == 'write' }
  read_op  <= valid_op { |o| o if o.type == 'read'  }
  fill_op  <= valid_op { |o| o if o.type == 'fill'  }
  trim_op  <= valid_op { |o| o if o.type == 'trim'  }
  seal_op  <= valid_op { |o| o if o.type == 'seal'  }
end
\end{lstlisting}

@@

\begin{lstlisting}[caption={Write}, label=lst:write]
bloom :write do
  temp :valid_write <= write_op.notin(found_op)
  log <+ valid_write{|o| [o.pos, 'valid', o.data]}
  ret <= valid_write{|o| [o.type, o.pos, o.epoch, 'ok']}
  ret <= write_op.notin(valid_write) { |o|
    [o.type, o.pos, o.epoch, 'read-only']
  }
end
\end{lstlisting}

@@

\begin{lstlisting}[caption={Read}, label=lst:read]
bloom :read do
  temp :valid_read <= (read_op * log).pairs(pos => pos) { |o, l|
    [o.type, o.pos, o.epoch, l.data] unless
      ['filled', 'trimmed'].include?(l.state)
  }
  ret <= valid_read { |e| [e.type, e.pos, e.epoch, e.data] }
  ret <= read_op.notin(valid_read, type => type) do |o|
    [o.type, o.pos, o.epoch, 'invalid']
  end
end
\end{lstlisting}

@@

As we have seen the subset of Bloom used in Brados is powerful enough to
express the semantics of the CORFU storage interface. We briefly sketch the
implementation of an additional interface found in Ceph and used by the
S3-compatible RADOS-Gateway service.

Channel
C++ code comparison

The subset of the Bloom language we are using in Brados is powerful enough to
express the CORFU language, as well as many of the object class interfaces that
are already included in Ceph. We briefly sketch the Bloom imlementation of an
object class that provides a locking service. This is an interesting example
because Bloom implementations use mechanisms for introducing points of
coordination in execution, which in the context of Ceph, may be written in
Bloom itself.

\begin{lstlisting}[caption={Hello}, label=lab]
class RefCount
  include Bud

  state do
    table :refs, [:tag]
    channel :remove_chn
    interface input, :get_op, [:tag]
    interface input, :put_op, [:tag]
    interface output, :ret, [:status]
  end

@@

    temp :remove <= found.notin(refs)	
    remove_chn <~ remove
  end
end
\end{lstlisting}

@@

Include performance management topic in future work.

>>>>

\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\lstset{frame=tb,
  language=Ruby,
  basicstyle={\scriptsize\ttfamily},
  breaklines=true,
  captionpos=b,
  numbers=left,
  numbersep=5pt,
  numberstyle=\tiny\color{mygray},
  xleftmargin=2em,
frame=l,
framexleftmargin=1.5em
}

%  aboveskip=3mm,
%  belowskip=3mm,
%  showstringspaces=false,
%  %columns=flexible,
%  basicstyle={\scriptsize\ttfamily},
%  %numberstyle=\tiny\color{gray},
%  %keywordstyle=\color{blue},
%  %commentstyle=\color{dkgreen},
%  %stringstyle=\color{mauve},
%  %breaklines=true,
%  breakatwhitespace=true,
%  tabsize=3
%}

@@

\begin{lstlisting}[caption={State Declaration}, label=lst:state]
state do
  table :epoch, [:epoch]
  table :log, [:pos] => [:state, :data]

  interface input, :op,
    [:type, :pos, :epoch] => [:data]

  interface output, :ret,
    [:type, :pos, :epoch] => [:retval]

@@

  # op's position was or was not found in the log
  scratch :found_op,    op.schema
  scratch :notfound_op, op.schema
end
\end{lstlisting}

\begin{lstlisting}[caption={Setup}, label=lst:setup]
bloom do
  # op's position found in log
  found_op <= (valid_op * log).lefts(pos => pos)
  notfound_op <= valid_op.notin(found_op)

  # demux on operation type
  write_op <= valid_op {|o| o if o.type == 'write'}
  read_op  <= valid_op {|o| o if o.type == 'read'}
  fill_op  <= valid_op {|o| o if o.type == 'fill'}
  trim_op  <= valid_op {|o| o if o.type == 'trim'}
  seal_op  <= valid_op {|o| o if o.type == 'seal'}
end
\end{lstlisting}

@@

\begin{lstlisting}[caption={Write}, label=lst:write]
bloom :write do
  temp :valid_write <= write_op.notin(found_op)
  log <+ valid_write{ |o| [o.pos, 'valid', o.data]}
  ret <= valid_write{ |o|
    [o.type, o.pos, o.epoch, 'ok'] }
  ret <= write_op.notin(valid_write) {|o|
    [o.type, o.pos, o.epoch, 'read-only'] }
end
\end{lstlisting}

@@

\begin{lstlisting}[caption={Read}, label=lst:read]
bloom :read do
  temp :ok_read <= (read_op * log).pairs(pos => pos) { |o, l|
    [o.type, o.pos, o.epoch, l.data] unless
      ['filled', 'trimmed'].include?(l.state) }
  ret <= ok_read { |e|
    [e.type, e.pos, e.epoch, e.data] }
  ret <= read_op.notin(ok_read, type=>type) do |o|
    [o.type, o.pos, o.epoch, 'invalid']
  end
end
\end{lstlisting}

@@

As we have seen the subset of Bloom used in Brados is powerful enough to
express the semantics of the CORFU storage interface. We briefly sketch the
implementation of an additional interface found in Ceph that used by the
S3-compatible RADOS-Gateway service to perform object reference counting.

The reference count implementation that we describe does not use a typical
counter, but rather uses a tag-based interface in which each \emph{reference}
to an object is identified by a application-level tag. Each tag represents
one or more references allowing an idempotent property useful for application
that may replay reference count operations out of order. Otherwise the
semantics are straightforward: clients record references to an object, and
when the number of references drop to zero the object is removed.

The reference implementation in Ceph is 300 LOC written in C++. The version
written in Brados is shown in Listing~\ref{lst:rc}.

 Unlike the CORFU interface
that took advantage of a mapping between backend storage and collections, the
reference counting interface requires access to RADOS runtime facilities that
allow objects to be removed. This is modeled in Brados using the Bloom
\emph{channel} abstraction that is used to transmit facts between entities in
the system; we effectively send a message to the runtime with our removal
request.

\begin{lstlisting}[caption={Reference counting interface}, label=lst:rc]
state do
  table :refs, [:tag]
  channel :remove_chn
  interface input, :get_op, [:tag]
  interface input, :put_op, [:tag]
  interface output, :ret, [:status]
end

@@

  temp :remove <= found.notin(refs)	
  remove_chn <~ remove
end
\end{lstlisting}

@@

Much work remains to be done.  While we demonstrated some of the immediate
benefits of separating logical and physical specifications of component
compositions, the ``dream’’ of data independence is to take the physical
design question completely off the table for programmers.  We are currently
exploring both cost-based optimizers and autotuners to automate the selection
of a high performance physical design consistent with a given logical
specification.

The database literature on materialized view maintenance and selection can be
brought to bear in the context of programmable storage as well.  Recall the
sequencer logic of Corfu described in Section XX.  The Bloom specification
(Figure YY) simply states that the current value of the sequencer is the
result of evaluating an aggregate query over the log relation.  This means
that the system may choose any strategy to maintain a sequence value that is
consistent with this constraint.  The physical design decision to store and
update a memory-resident sequence value rather re-computing the maximal value
at each request is a view materialization decision that could similarly be
made by a compile-time optimizer.

The future of programmable storage is likely to witness a further
decomposition of services to allow greater flexibility in customizing
deployments.  More than likely, not all of these services will assume or
provide strongly-consistent operations by default.  Instead, programmers will
have the flexibility (and burden!) of trading off between performance (offered
by weakly consistent subsystems) and programmability.  Languages such as
Bloom---which provide static analyses to determine when loosely-ordered
operations nevertheless produce consistent outcomes---will become increasingly
relevant in this domain.

%Include performance management topic in future work.
